---
title: "Pilot Retention Subfactor Analysis"
subtitle: "Phase 1A: Within-Construct Sub-Item Rankings and Demographic Comparisons"
author: "Analysis Team"
date: today
format:
  docx:
    toc: true
    number-sections: true
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    theme: cosmo
    number-sections: true
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This analysis examines the 31 sub-item rankings within each retention construct and compares them across demographics.

# Setup

```{r setup}
library(tidyverse)
library(rstatix)
library(knitr)
library(janitor)
library(DescTools)

options(scipen = 999)
knitr::opts_chunk$set(fig.width = 10, fig.height = 6)

# Create output directories
dir.create("../../output/tables/subfactor_analysis", showWarnings = FALSE, recursive = TRUE)
dir.create("../../output/figures/subfactor_analysis", showWarnings = FALSE, recursive = TRUE)

# Helper function to safely run Mann-Whitney U tests with effect sizes
safe_wilcox_test <- function(data, formula, exact = FALSE) {
  # Check if we have at least 2 groups with data
  group_var <- all.vars(formula)[2]
  n_groups <- data %>%
    filter(!is.na(!!sym(group_var))) %>%
    pull(!!sym(group_var)) %>%
    n_distinct()

  if (n_groups < 2) {
    # Return empty result if not enough groups
    return(tibble(
      .y. = character(),
      group1 = character(),
      group2 = character(),
      n1 = integer(),
      n2 = integer(),
      statistic = numeric(),
      p = numeric(),
      effsize = numeric(),
      magnitude = character()
    ))
  }

  # Run the test (this should succeed for all valid items)
  test_result <- tryCatch({
    wilcox_test(data, formula, exact = exact)
  },
  error = function(e) {
    # Return empty on test failure
    return(tibble(
      .y. = character(),
      group1 = character(),
      group2 = character(),
      n1 = integer(),
      n2 = integer(),
      statistic = numeric(),
      p = numeric()
    ))
  })

  # If test failed, return empty
  if (nrow(test_result) == 0) {
    return(tibble(
      .y. = character(),
      group1 = character(),
      group2 = character(),
      n1 = integer(),
      n2 = integer(),
      statistic = numeric(),
      p = numeric(),
      effsize = numeric(),
      magnitude = character()
    ))
  }

  # Try to compute effect sizes (may fail for sparse items)
  effsize_result <- tryCatch({
    wilcox_effsize(data, formula)
  },
  error = function(e) {
    # Return NA effect sizes for items that fail
    test_result %>%
      select(.y.) %>%
      distinct() %>%
      mutate(
        effsize = NA_real_,
        magnitude = NA_character_
      )
  })

  # Combine results with proper join
  test_result %>%
    left_join(
      effsize_result %>% select(.y., effsize, magnitude),
      by = ".y."
    )
}

# Helper function to safely run Kruskal-Wallis tests with effect sizes
safe_kruskal_test <- function(data, formula) {
  # Check if we have at least 2 groups with data
  group_var <- all.vars(formula)[2]
  n_groups <- data %>%
    filter(!is.na(!!sym(group_var))) %>%
    pull(!!sym(group_var)) %>%
    n_distinct()

  if (n_groups < 2) {
    # Return empty result if not enough groups
    return(tibble(
      .y. = character(),
      n = integer(),
      statistic = numeric(),
      df = integer(),
      p = numeric(),
      effsize = numeric(),
      magnitude = character()
    ))
  }

  # Run the test (this should succeed for all valid items)
  test_result <- tryCatch({
    kruskal_test(data, formula)
  },
  error = function(e) {
    # Return empty on test failure
    return(tibble(
      .y. = character(),
      n = integer(),
      statistic = numeric(),
      df = integer(),
      p = numeric()
    ))
  })

  # If test failed, return empty
  if (nrow(test_result) == 0) {
    return(tibble(
      .y. = character(),
      n = integer(),
      statistic = numeric(),
      df = integer(),
      p = numeric(),
      effsize = numeric(),
      magnitude = character()
    ))
  }

  # Try to compute effect sizes (may fail for sparse items)
  effsize_result <- tryCatch({
    kruskal_effsize(data, formula)
  },
  error = function(e) {
    # Return NA effect sizes for items that fail
    test_result %>%
      select(.y.) %>%
      distinct() %>%
      mutate(
        effsize = NA_real_,
        magnitude = NA_character_
      )
  })

  # Combine results with proper join
  test_result %>%
    left_join(
      effsize_result %>% select(.y., effsize, magnitude),
      by = ".y."
    )
}

# Variable labels from survey instrument
financial_labels <- c(
  "financial_1" = "Competitive salary",
  "financial_2" = "Allowances and soft pay",
  "financial_3" = "Benefits package",
  "financial_4" = "Disability insurance",
  "financial_5" = "Job security",
  "financial_6" = "New hire/longevity bonuses"
)

qol_labels <- c(
  "qo_l_1" = "Predictable schedule",
  "qo_l_2" = "Vacation time",
  "qo_l_3" = "Family-friendly policies",
  "qo_l_4" = "Home base possibility",
  "qo_l_5" = "Travel benefits",
  "qo_l_6" = "Work-life balance"
)

professional_labels <- c(
  "professional_1" = "Financially stable airline",
  "professional_2" = "Rapid upgrade opportunity",
  "professional_3" = "Fly larger aircraft",
  "professional_4" = "Merit-based promotion",
  "professional_5" = "Seniority-based upgrade"
)

recognition_labels <- c(
  "recognition_1" = "Professional uniforms",
  "recognition_3" = "Vacation for long service",
  "recognition_4" = "Professional respect/recognition",
  "recognition_7" = "Airline recognition"
)

schedule_labels <- c(
  "schedule_1" = "Fixed schedule",
  "schedule_2" = "Variable schedule",
  "schedule_3" = "Flexible work rules",
  "schedule_4" = "Bid line seniority system",
  "schedule_5" = "Vacation bidding rules"
)

operational_labels <- c(
  "operational_1" = "Unambiguous SOPs",
  "operational_2" = "Proactive training",
  "operational_3" = "Well-maintained aircraft",
  "operational_4" = "Well-equipped aircraft",
  "operational_5" = "Highly skilled pilots"
)
```

# Data Loading

```{r load-data}
# Load data (same as retention.Rmd)
raw <- read_csv("../../data/processed/retention.csv", show_col_types = FALSE)

retention_data <- raw %>%
  slice(-(1:2)) %>%
  clean_names()

# Filter to valid responses
included_data <- retention_data %>%
  mutate(
    finished = toupper(as.character(finished)),
    informed_consent = toupper(as.character(informed_consent)),
    screener = toupper(as.character(screener))
  ) %>%
  filter(
    finished == "TRUE",
    informed_consent == "YES",
    screener == "YES"
  )

cat("Valid observations:", nrow(included_data), "\n")
```

# Demographic Groups

```{r demographics}
# Create demographic groups (matching retention.Rmd approach)
included_data <- included_data %>%
  mutate(
    # Numeric conversions
    age = as.numeric(age),
    experience = as.numeric(experience),

    # Age groups
    age_group = if_else(age <= 35, "≤35", ">35"),

    # Experience quartiles
    exp_q = ntile(experience, 4),
    exp_q = factor(exp_q, labels = c("Q1", "Q2", "Q3", "Q4")),

    # Gender
    gender_clean = case_when(
      tolower(gender) %in% c("male", "m") ~ "Male",
      tolower(gender) %in% c("female", "f") ~ "Female",
      TRUE ~ NA_character_
    ),

    # Position
    position = position_position,

    # Military
    military_clean = case_when(
      str_detect(tolower(military_experience), "currently serve|previously served") ~ "Military",
      str_detect(tolower(military_experience), "never") ~ "Civilian",
      TRUE ~ NA_character_
    )
  )

# Summary
cat("\n=== DEMOGRAPHICS ===\n")
cat("Age groups:\n"); table(included_data$age_group) %>% print()
cat("\nGender:\n"); table(included_data$gender_clean, useNA = "ifany") %>% print()
cat("\nPosition:\n"); table(included_data$position, useNA = "ifany") %>% print()
cat("\nExperience:\n"); table(included_data$exp_q) %>% print()
cat("\nMilitary:\n"); table(included_data$military_clean, useNA = "ifany") %>% print()
```

# Financial Subfactors

```{r financial}
cat("\n========================================\n")
cat("FINANCIAL SUBFACTORS\n")
cat("========================================\n\n")

# Convert to long format
financial_long <- included_data %>%
  select(age_group, gender_clean, position, exp_q, military_clean,
         financial_1:financial_6) %>%
  mutate(respondent_id = row_number()) %>%  # Create persistent ID before pivoting
  mutate(across(financial_1:financial_6, as.numeric)) %>%
  pivot_longer(
    cols = financial_1:financial_6,
    names_to = "item",
    values_to = "rank"
  ) %>%
  filter(!is.na(rank))

# Descriptive stats
financial_desc <- financial_long %>%
  group_by(item) %>%
  summarise(
    n = n(),
    median = median(rank),
    mean = mean(rank),
    sd = sd(rank),
    pct_rank1 = sum(rank == 1) / n() * 100
  ) %>%
  mutate(item = recode(item, !!!financial_labels))

kable(financial_desc, digits = 2, caption = "Financial Subfactors - Descriptive Statistics")

# Friedman test
friedman_financial <- financial_long %>%
  friedman_test(rank ~ item | respondent_id)

kable(friedman_financial, digits = 3, caption = "Financial: Friedman Test (Within-Construct Comparison)")

# Median rank visualization
financial_plot <- financial_desc %>%
  mutate(item = fct_reorder(item, desc(median))) %>%
  ggplot(aes(x = median, y = item)) +
  geom_col(fill = "steelblue", alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f", median)), hjust = -0.2, size = 3) +
  labs(
    title = "Financial Subfactor Priorities",
    subtitle = "Median rank (lower = higher priority)",
    x = "Median Rank",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

print(financial_plot)
ggsave("../../output/figures/subfactor_analysis/financial_median_ranks.png",
       financial_plot, width = 8, height = 5, dpi = 300)

# Age group comparisons (Mann-Whitney U)
mw_age_fin <- financial_long %>%
  filter(!is.na(age_group)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ age_group, exact = FALSE) %>%
  ungroup()

if (nrow(mw_age_fin) > 0) {
  mw_age_fin <- mw_age_fin %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!financial_labels))
}

kable(mw_age_fin, digits = 3, caption = "Financial: Age Group Comparisons (Mann-Whitney U)")

# Gender comparisons
mw_gender_fin <- financial_long %>%
  filter(!is.na(gender_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ gender_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_gender_fin) > 0) {
  mw_gender_fin <- mw_gender_fin %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!financial_labels))
}

kable(mw_gender_fin, digits = 3, caption = "Financial: Gender Comparisons (Mann-Whitney U)")

# Position comparisons
mw_position_fin <- financial_long %>%
  filter(!is.na(position)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ position, exact = FALSE) %>%
  ungroup()

if (nrow(mw_position_fin) > 0) {
  mw_position_fin <- mw_position_fin %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!financial_labels))
}

kable(mw_position_fin, digits = 3, caption = "Financial: Position Comparisons (Mann-Whitney U)")

# Military comparisons
mw_military_fin <- financial_long %>%
  filter(!is.na(military_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ military_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_military_fin) > 0) {
  mw_military_fin <- mw_military_fin %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!financial_labels))
}

kable(mw_military_fin, digits = 3, caption = "Financial: Military Comparisons (Mann-Whitney U)")

# Experience quartile comparisons (Kruskal-Wallis)
kw_exp_fin <- financial_long %>%
  filter(!is.na(exp_q)) %>%
  group_by(item) %>%
  safe_kruskal_test(rank ~ exp_q) %>%
  ungroup()

if (nrow(kw_exp_fin) > 0) {
  kw_exp_fin <- kw_exp_fin %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, df, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!financial_labels))
}

kable(kw_exp_fin, digits = 3, caption = "Financial: Experience Comparisons (Kruskal-Wallis)")

# Save tables to CSV
write_csv(financial_desc, "../../output/tables/subfactor_analysis/financial_descriptive.csv")
write_csv(friedman_financial, "../../output/tables/subfactor_analysis/financial_friedman.csv")
if (nrow(mw_age_fin) > 0) write_csv(mw_age_fin, "../../output/tables/subfactor_analysis/financial_age.csv")
if (nrow(mw_gender_fin) > 0) write_csv(mw_gender_fin, "../../output/tables/subfactor_analysis/financial_gender.csv")
if (nrow(mw_position_fin) > 0) write_csv(mw_position_fin, "../../output/tables/subfactor_analysis/financial_position.csv")
if (nrow(mw_military_fin) > 0) write_csv(mw_military_fin, "../../output/tables/subfactor_analysis/financial_military.csv")
if (nrow(kw_exp_fin) > 0) write_csv(kw_exp_fin, "../../output/tables/subfactor_analysis/financial_experience.csv")
```

# Quality of Life Subfactors

```{r qol}
cat("\n========================================\n")
cat("QUALITY OF LIFE SUBFACTORS\n")
cat("========================================\n\n")

# Convert to long format
qol_long <- included_data %>%
  select(age_group, gender_clean, position, exp_q, military_clean,
         qo_l_1:qo_l_6) %>%
  mutate(respondent_id = row_number()) %>%  # Create persistent ID before pivoting
  mutate(across(qo_l_1:qo_l_6, as.numeric)) %>%
  pivot_longer(
    cols = qo_l_1:qo_l_6,
    names_to = "item",
    values_to = "rank"
  ) %>%
  filter(!is.na(rank))

# Descriptive stats
qol_desc <- qol_long %>%
  group_by(item) %>%
  summarise(
    n = n(),
    median = median(rank),
    mean = mean(rank),
    sd = sd(rank),
    pct_rank1 = sum(rank == 1) / n() * 100
  ) %>%
  mutate(item = recode(item, !!!qol_labels))

kable(qol_desc, digits = 2, caption = "QoL Subfactors - Descriptive Statistics")

# Friedman test
friedman_qol <- qol_long %>%
  friedman_test(rank ~ item | respondent_id)

kable(friedman_qol, digits = 3, caption = "QoL: Friedman Test (Within-Construct Comparison)")

# Median rank visualization
qol_plot <- qol_desc %>%
  mutate(item = fct_reorder(item, desc(median))) %>%
  ggplot(aes(x = median, y = item)) +
  geom_col(fill = "forestgreen", alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f", median)), hjust = -0.2, size = 3) +
  labs(
    title = "Quality of Life Subfactor Priorities",
    subtitle = "Median rank (lower = higher priority)",
    x = "Median Rank",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

print(qol_plot)
ggsave("../../output/figures/subfactor_analysis/qol_median_ranks.png",
       qol_plot, width = 8, height = 5, dpi = 300)

# Age group
mw_age_qol <- qol_long %>%
  filter(!is.na(age_group)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ age_group, exact = FALSE) %>%
  ungroup()

if (nrow(mw_age_qol) > 0) {
  mw_age_qol <- mw_age_qol %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!qol_labels))
}

kable(mw_age_qol, digits = 3, caption = "QoL: Age Group Comparisons")

# Gender
mw_gender_qol <- qol_long %>%
  filter(!is.na(gender_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ gender_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_gender_qol) > 0) {
  mw_gender_qol <- mw_gender_qol %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!qol_labels))
}

kable(mw_gender_qol, digits = 3, caption = "QoL: Gender Comparisons")

# Position
mw_position_qol <- qol_long %>%
  filter(!is.na(position)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ position, exact = FALSE) %>%
  ungroup()

if (nrow(mw_position_qol) > 0) {
  mw_position_qol <- mw_position_qol %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!qol_labels))
}

kable(mw_position_qol, digits = 3, caption = "QoL: Position Comparisons")

# Military
mw_military_qol <- qol_long %>%
  filter(!is.na(military_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ military_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_military_qol) > 0) {
  mw_military_qol <- mw_military_qol %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!qol_labels))
}

kable(mw_military_qol, digits = 3, caption = "QoL: Military Comparisons")

# Experience
kw_exp_qol <- qol_long %>%
  filter(!is.na(exp_q)) %>%
  group_by(item) %>%
  safe_kruskal_test(rank ~ exp_q) %>%
  ungroup()

if (nrow(kw_exp_qol) > 0) {
  kw_exp_qol <- kw_exp_qol %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, df, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!qol_labels))
}

kable(kw_exp_qol, digits = 3, caption = "QoL: Experience Comparisons")

# Save tables to CSV
write_csv(qol_desc, "../../output/tables/subfactor_analysis/qol_descriptive.csv")
write_csv(friedman_qol, "../../output/tables/subfactor_analysis/qol_friedman.csv")
if (nrow(mw_age_qol) > 0) write_csv(mw_age_qol, "../../output/tables/subfactor_analysis/qol_age.csv")
if (nrow(mw_gender_qol) > 0) write_csv(mw_gender_qol, "../../output/tables/subfactor_analysis/qol_gender.csv")
if (nrow(mw_position_qol) > 0) write_csv(mw_position_qol, "../../output/tables/subfactor_analysis/qol_position.csv")
if (nrow(mw_military_qol) > 0) write_csv(mw_military_qol, "../../output/tables/subfactor_analysis/qol_military.csv")
if (nrow(kw_exp_qol) > 0) write_csv(kw_exp_qol, "../../output/tables/subfactor_analysis/qol_experience.csv")
```

# Professional Subfactors

```{r professional}
cat("\n========================================\n")
cat("PROFESSIONAL SUBFACTORS\n")
cat("========================================\n\n")

professional_long <- included_data %>%
  select(age_group, gender_clean, position, exp_q, military_clean,
         professional_1:professional_5) %>%
  mutate(respondent_id = row_number()) %>%  # Create persistent ID before pivoting
  mutate(across(professional_1:professional_5, as.numeric)) %>%
  pivot_longer(
    cols = professional_1:professional_5,
    names_to = "item",
    values_to = "rank"
  ) %>%
  filter(!is.na(rank))

# Descriptive
professional_desc <- professional_long %>%
  group_by(item) %>%
  summarise(n = n(), median = median(rank), mean = mean(rank),
            sd = sd(rank), pct_rank1 = sum(rank == 1) / n() * 100) %>%
  mutate(item = recode(item, !!!professional_labels))
kable(professional_desc, digits = 2, caption = "Professional Subfactors - Descriptive Statistics")

# Friedman
friedman_prof <- professional_long %>%
  friedman_test(rank ~ item | respondent_id)
kable(friedman_prof, digits = 3, caption = "Professional: Friedman Test (Within-Construct Comparison)")

# Median rank visualization
professional_plot <- professional_desc %>%
  mutate(item = fct_reorder(item, desc(median))) %>%
  ggplot(aes(x = median, y = item)) +
  geom_col(fill = "darkorange", alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f", median)), hjust = -0.2, size = 3) +
  labs(
    title = "Professional Opportunity Subfactor Priorities",
    subtitle = "Median rank (lower = higher priority)",
    x = "Median Rank",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

print(professional_plot)
ggsave("../../output/figures/subfactor_analysis/professional_median_ranks.png",
       professional_plot, width = 8, height = 5, dpi = 300)

# Demographics
mw_age_prof <- professional_long %>%
  filter(!is.na(age_group)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ age_group, exact = FALSE) %>%
  ungroup()

if (nrow(mw_age_prof) > 0) {
  mw_age_prof <- mw_age_prof %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!professional_labels))
}

kable(mw_age_prof, digits = 3, caption = "Professional: Age Comparisons")

mw_gender_prof <- professional_long %>%
  filter(!is.na(gender_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ gender_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_gender_prof) > 0) {
  mw_gender_prof <- mw_gender_prof %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!professional_labels))
}

kable(mw_gender_prof, digits = 3, caption = "Professional: Gender Comparisons")

mw_position_prof <- professional_long %>%
  filter(!is.na(position)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ position, exact = FALSE) %>%
  ungroup()

if (nrow(mw_position_prof) > 0) {
  mw_position_prof <- mw_position_prof %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!professional_labels))
}

kable(mw_position_prof, digits = 3, caption = "Professional: Position Comparisons")

mw_military_prof <- professional_long %>%
  filter(!is.na(military_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ military_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_military_prof) > 0) {
  mw_military_prof <- mw_military_prof %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!professional_labels))
}

kable(mw_military_prof, digits = 3, caption = "Professional: Military Comparisons")

kw_exp_prof <- professional_long %>%
  filter(!is.na(exp_q)) %>%
  group_by(item) %>%
  safe_kruskal_test(rank ~ exp_q) %>%
  ungroup()

if (nrow(kw_exp_prof) > 0) {
  kw_exp_prof <- kw_exp_prof %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, df, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!professional_labels))
}

kable(kw_exp_prof, digits = 3, caption = "Professional: Experience Comparisons")

# Save tables to CSV
write_csv(professional_desc, "../../output/tables/subfactor_analysis/professional_descriptive.csv")
write_csv(friedman_prof, "../../output/tables/subfactor_analysis/professional_friedman.csv")
if (nrow(mw_age_prof) > 0) write_csv(mw_age_prof, "../../output/tables/subfactor_analysis/professional_age.csv")
if (nrow(mw_gender_prof) > 0) write_csv(mw_gender_prof, "../../output/tables/subfactor_analysis/professional_gender.csv")
if (nrow(mw_position_prof) > 0) write_csv(mw_position_prof, "../../output/tables/subfactor_analysis/professional_position.csv")
if (nrow(mw_military_prof) > 0) write_csv(mw_military_prof, "../../output/tables/subfactor_analysis/professional_military.csv")
if (nrow(kw_exp_prof) > 0) write_csv(kw_exp_prof, "../../output/tables/subfactor_analysis/professional_experience.csv")
```

# Recognition Subfactors

```{r recognition}
cat("\n========================================\n")
cat("RECOGNITION SUBFACTORS\n")
cat("========================================\n\n")

recognition_long <- included_data %>%
  select(age_group, gender_clean, position, exp_q, military_clean,
         recognition_1, recognition_3, recognition_4, recognition_7) %>%
  mutate(respondent_id = row_number()) %>%  # Create persistent ID before pivoting
  mutate(across(starts_with("recognition"), as.numeric)) %>%
  pivot_longer(
    cols = starts_with("recognition"),
    names_to = "item",
    values_to = "rank"
  ) %>%
  filter(!is.na(rank))

# Descriptive
recognition_desc <- recognition_long %>%
  group_by(item) %>%
  summarise(n = n(), median = median(rank), mean = mean(rank),
            sd = sd(rank), pct_rank1 = sum(rank == 1) / n() * 100) %>%
  mutate(item = recode(item, !!!recognition_labels))
kable(recognition_desc, digits = 2, caption = "Recognition Subfactors - Descriptive Statistics")

# Friedman
friedman_recog <- recognition_long %>%
  friedman_test(rank ~ item | respondent_id)
kable(friedman_recog, digits = 3, caption = "Recognition: Friedman Test (Within-Construct Comparison)")

# Median rank visualization
recognition_plot <- recognition_desc %>%
  mutate(item = fct_reorder(item, desc(median))) %>%
  ggplot(aes(x = median, y = item)) +
  geom_col(fill = "purple", alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f", median)), hjust = -0.2, size = 3) +
  labs(
    title = "Recognition Subfactor Priorities",
    subtitle = "Median rank (lower = higher priority)",
    x = "Median Rank",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

print(recognition_plot)
ggsave("../../output/figures/subfactor_analysis/recognition_median_ranks.png",
       recognition_plot, width = 8, height = 5, dpi = 300)

# Demographics
mw_age_recog <- recognition_long %>%
  filter(!is.na(age_group)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ age_group, exact = FALSE) %>%
  ungroup()

if (nrow(mw_age_recog) > 0) {
  mw_age_recog <- mw_age_recog %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!recognition_labels))
}

kable(mw_age_recog, digits = 3, caption = "Recognition: Age Comparisons")

mw_gender_recog <- recognition_long %>%
  filter(!is.na(gender_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ gender_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_gender_recog) > 0) {
  mw_gender_recog <- mw_gender_recog %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!recognition_labels))
}

kable(mw_gender_recog, digits = 3, caption = "Recognition: Gender Comparisons")

mw_position_recog <- recognition_long %>%
  filter(!is.na(position)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ position, exact = FALSE) %>%
  ungroup()

if (nrow(mw_position_recog) > 0) {
  mw_position_recog <- mw_position_recog %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!recognition_labels))
}

kable(mw_position_recog, digits = 3, caption = "Recognition: Position Comparisons")

mw_military_recog <- recognition_long %>%
  filter(!is.na(military_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ military_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_military_recog) > 0) {
  mw_military_recog <- mw_military_recog %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!recognition_labels))
}

kable(mw_military_recog, digits = 3, caption = "Recognition: Military Comparisons")

kw_exp_recog <- recognition_long %>%
  filter(!is.na(exp_q)) %>%
  group_by(item) %>%
  safe_kruskal_test(rank ~ exp_q) %>%
  ungroup()

if (nrow(kw_exp_recog) > 0) {
  kw_exp_recog <- kw_exp_recog %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, df, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!recognition_labels))
}

kable(kw_exp_recog, digits = 3, caption = "Recognition: Experience Comparisons")

# Save tables to CSV
write_csv(recognition_desc, "../../output/tables/subfactor_analysis/recognition_descriptive.csv")
write_csv(friedman_recog, "../../output/tables/subfactor_analysis/recognition_friedman.csv")
if (nrow(mw_age_recog) > 0) write_csv(mw_age_recog, "../../output/tables/subfactor_analysis/recognition_age.csv")
if (nrow(mw_gender_recog) > 0) write_csv(mw_gender_recog, "../../output/tables/subfactor_analysis/recognition_gender.csv")
if (nrow(mw_position_recog) > 0) write_csv(mw_position_recog, "../../output/tables/subfactor_analysis/recognition_position.csv")
if (nrow(mw_military_recog) > 0) write_csv(mw_military_recog, "../../output/tables/subfactor_analysis/recognition_military.csv")
if (nrow(kw_exp_recog) > 0) write_csv(kw_exp_recog, "../../output/tables/subfactor_analysis/recognition_experience.csv")
```

# Schedule Subfactors

```{r schedule}
cat("\n========================================\n")
cat("SCHEDULE SUBFACTORS\n")
cat("========================================\n\n")

schedule_long <- included_data %>%
  select(age_group, gender_clean, position, exp_q, military_clean,
         schedule_1:schedule_5) %>%
  mutate(respondent_id = row_number()) %>%  # Create persistent ID before pivoting
  mutate(across(schedule_1:schedule_5, as.numeric)) %>%
  pivot_longer(
    cols = schedule_1:schedule_5,
    names_to = "item",
    values_to = "rank"
  ) %>%
  filter(!is.na(rank))

# Descriptive
schedule_desc <- schedule_long %>%
  group_by(item) %>%
  summarise(n = n(), median = median(rank), mean = mean(rank),
            sd = sd(rank), pct_rank1 = sum(rank == 1) / n() * 100) %>%
  mutate(item = recode(item, !!!schedule_labels))
kable(schedule_desc, digits = 2, caption = "Schedule Subfactors - Descriptive Statistics")

# Friedman
friedman_sched <- schedule_long %>%
  friedman_test(rank ~ item | respondent_id)
kable(friedman_sched, digits = 3, caption = "Schedule: Friedman Test (Within-Construct Comparison)")

# Median rank visualization
schedule_plot <- schedule_desc %>%
  mutate(item = fct_reorder(item, desc(median))) %>%
  ggplot(aes(x = median, y = item)) +
  geom_col(fill = "firebrick", alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f", median)), hjust = -0.2, size = 3) +
  labs(
    title = "Schedule Subfactor Priorities",
    subtitle = "Median rank (lower = higher priority)",
    x = "Median Rank",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

print(schedule_plot)
ggsave("../../output/figures/subfactor_analysis/schedule_median_ranks.png",
       schedule_plot, width = 8, height = 5, dpi = 300)

# Demographics
mw_age_sched <- schedule_long %>%
  filter(!is.na(age_group)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ age_group, exact = FALSE) %>%
  ungroup()

if (nrow(mw_age_sched) > 0) {
  mw_age_sched <- mw_age_sched %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!schedule_labels))
}

kable(mw_age_sched, digits = 3, caption = "Schedule: Age Comparisons")

mw_gender_sched <- schedule_long %>%
  filter(!is.na(gender_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ gender_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_gender_sched) > 0) {
  mw_gender_sched <- mw_gender_sched %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!schedule_labels))
}

kable(mw_gender_sched, digits = 3, caption = "Schedule: Gender Comparisons")

mw_position_sched <- schedule_long %>%
  filter(!is.na(position)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ position, exact = FALSE) %>%
  ungroup()

if (nrow(mw_position_sched) > 0) {
  mw_position_sched <- mw_position_sched %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!schedule_labels))
}

kable(mw_position_sched, digits = 3, caption = "Schedule: Position Comparisons")

mw_military_sched <- schedule_long %>%
  filter(!is.na(military_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ military_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_military_sched) > 0) {
  mw_military_sched <- mw_military_sched %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!schedule_labels))
}

kable(mw_military_sched, digits = 3, caption = "Schedule: Military Comparisons")

kw_exp_sched <- schedule_long %>%
  filter(!is.na(exp_q)) %>%
  group_by(item) %>%
  safe_kruskal_test(rank ~ exp_q) %>%
  ungroup()

if (nrow(kw_exp_sched) > 0) {
  kw_exp_sched <- kw_exp_sched %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, df, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!schedule_labels))
}

kable(kw_exp_sched, digits = 3, caption = "Schedule: Experience Comparisons")

# Save tables to CSV
write_csv(schedule_desc, "../../output/tables/subfactor_analysis/schedule_descriptive.csv")
write_csv(friedman_sched, "../../output/tables/subfactor_analysis/schedule_friedman.csv")
if (nrow(mw_age_sched) > 0) write_csv(mw_age_sched, "../../output/tables/subfactor_analysis/schedule_age.csv")
if (nrow(mw_gender_sched) > 0) write_csv(mw_gender_sched, "../../output/tables/subfactor_analysis/schedule_gender.csv")
if (nrow(mw_position_sched) > 0) write_csv(mw_position_sched, "../../output/tables/subfactor_analysis/schedule_position.csv")
if (nrow(mw_military_sched) > 0) write_csv(mw_military_sched, "../../output/tables/subfactor_analysis/schedule_military.csv")
if (nrow(kw_exp_sched) > 0) write_csv(kw_exp_sched, "../../output/tables/subfactor_analysis/schedule_experience.csv")
```

# Operational Subfactors

```{r operational}
cat("\n========================================\n")
cat("OPERATIONAL SUBFACTORS\n")
cat("========================================\n\n")

operational_long <- included_data %>%
  select(age_group, gender_clean, position, exp_q, military_clean,
         operational_1:operational_5) %>%
  mutate(respondent_id = row_number()) %>%  # Create persistent ID before pivoting
  mutate(across(operational_1:operational_5, as.numeric)) %>%
  pivot_longer(
    cols = operational_1:operational_5,
    names_to = "item",
    values_to = "rank"
  ) %>%
  filter(!is.na(rank))

# Descriptive
operational_desc <- operational_long %>%
  group_by(item) %>%
  summarise(n = n(), median = median(rank), mean = mean(rank),
            sd = sd(rank), pct_rank1 = sum(rank == 1) / n() * 100) %>%
  mutate(item = recode(item, !!!operational_labels))
kable(operational_desc, digits = 2, caption = "Operational Subfactors - Descriptive Statistics")

# Friedman
friedman_oper <- operational_long %>%
  friedman_test(rank ~ item | respondent_id)
kable(friedman_oper, digits = 3, caption = "Operational: Friedman Test (Within-Construct Comparison)")

# Median rank visualization
operational_plot <- operational_desc %>%
  mutate(item = fct_reorder(item, desc(median))) %>%
  ggplot(aes(x = median, y = item)) +
  geom_col(fill = "darkslategray", alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f", median)), hjust = -0.2, size = 3) +
  labs(
    title = "Operational Subfactor Priorities",
    subtitle = "Median rank (lower = higher priority)",
    x = "Median Rank",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

print(operational_plot)
ggsave("../../output/figures/subfactor_analysis/operational_median_ranks.png",
       operational_plot, width = 8, height = 5, dpi = 300)

# Demographics
mw_age_oper <- operational_long %>%
  filter(!is.na(age_group)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ age_group, exact = FALSE) %>%
  ungroup()

if (nrow(mw_age_oper) > 0) {
  mw_age_oper <- mw_age_oper %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!operational_labels))
}

kable(mw_age_oper, digits = 3, caption = "Operational: Age Comparisons")

mw_gender_oper <- operational_long %>%
  filter(!is.na(gender_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ gender_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_gender_oper) > 0) {
  mw_gender_oper <- mw_gender_oper %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!operational_labels))
}

kable(mw_gender_oper, digits = 3, caption = "Operational: Gender Comparisons")

mw_position_oper <- operational_long %>%
  filter(!is.na(position)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ position, exact = FALSE) %>%
  ungroup()

if (nrow(mw_position_oper) > 0) {
  mw_position_oper <- mw_position_oper %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!operational_labels))
}

kable(mw_position_oper, digits = 3, caption = "Operational: Position Comparisons")

mw_military_oper <- operational_long %>%
  filter(!is.na(military_clean)) %>%
  group_by(item) %>%
  safe_wilcox_test(rank ~ military_clean, exact = FALSE) %>%
  ungroup()

if (nrow(mw_military_oper) > 0) {
  mw_military_oper <- mw_military_oper %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!operational_labels))
}

kable(mw_military_oper, digits = 3, caption = "Operational: Military Comparisons")

kw_exp_oper <- operational_long %>%
  filter(!is.na(exp_q)) %>%
  group_by(item) %>%
  safe_kruskal_test(rank ~ exp_q) %>%
  ungroup()

if (nrow(kw_exp_oper) > 0) {
  kw_exp_oper <- kw_exp_oper %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance("p.adj") %>%
    select(item, statistic, df, p, p.adj, p.adj.signif, effsize, magnitude) %>%
    mutate(item = recode(item, !!!operational_labels))
}

kable(kw_exp_oper, digits = 3, caption = "Operational: Experience Comparisons")

# Save tables to CSV
write_csv(operational_desc, "../../output/tables/subfactor_analysis/operational_descriptive.csv")
write_csv(friedman_oper, "../../output/tables/subfactor_analysis/operational_friedman.csv")
if (nrow(mw_age_oper) > 0) write_csv(mw_age_oper, "../../output/tables/subfactor_analysis/operational_age.csv")
if (nrow(mw_gender_oper) > 0) write_csv(mw_gender_oper, "../../output/tables/subfactor_analysis/operational_gender.csv")
if (nrow(mw_position_oper) > 0) write_csv(mw_position_oper, "../../output/tables/subfactor_analysis/operational_position.csv")
if (nrow(mw_military_oper) > 0) write_csv(mw_military_oper, "../../output/tables/subfactor_analysis/operational_military.csv")
if (nrow(kw_exp_oper) > 0) write_csv(kw_exp_oper, "../../output/tables/subfactor_analysis/operational_experience.csv")
```

# Conclusion

This analysis examined 31 subfactor rankings across six retention constructs (Financial, Quality of Life, Professional Opportunity, Recognition, Schedule, and Operational) and compared them across five demographic groups: age (≤35 vs >35), gender, position (Captain vs First Officer), military experience, and years of experience (quartiles).

**Statistical Methods:** Mann-Whitney U tests were used for binary demographic comparisons, Kruskal-Wallis tests for experience quartiles, and Friedman tests for within-construct comparisons. All tests used proper respondent identifiers to maintain the repeated-measures structure of the data.

**Key Finding:** After conducting 155 demographic comparison tests with Bonferroni correction for multiple comparisons, **no significant demographic differences were detected** (p.adj < 0.05) in how pilots ranked subfactor priorities within any retention construct.

Regardless of age, gender, position, military background, or years of experience, pilots demonstrate similar preferences for specific retention factors within each construct.

# Session Info

```{r session}
sessionInfo()
```
